TODO: Title

### 2013-06-26 Doc/Media2013/media2013.tex
preparing MEDIA paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@597 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

% http://www.miccai2013.org/submission_guideline.html
\documentclass{llncs}

\usepackage{amsmath}
\usepackage{makeidx}
\usepackage[pdftex]{graphicx}
\usepackage{subfigure}

\def\wrt{w.\,r.\,t.}
\def\eg{e.\,g.}
\def\ie{i.\,e.}
\def\Dash{\nobreak\,---\penalty-500\,}

@@

\renewcommand{\Vec}[1]{\mathbf{#1}}
\newcommand{\Mat}[1]{\mathbf{#1}}

\begin{document}
%
%\frontmatter          % for the preliminaries
%
\mainmatter              % start of the contributions
%
\title{Complete Real-Time Liver Model Including Glisson's Capsule, Vascularization and Parenchyma}
%
\titlerunning{Complete Liver Model}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Anonymous}
%\author{%
%Tom\'a\v{s} Golembiovsk\'y\inst{1,2} \and%
%Igor Peterl\'ik\inst{3} \and%
%Christian Duriez\inst{2,4} \and%
%Stephan\'e Cotin\inst{2}%
%}
%
%\authorrunning{Tom\'a\vs Golembiovsk\'y et al.} % abbreviated author list (for running head)
%
%\institute{%
%Faculty of Informatics, Masaryk University, Brno, Czech Republic\and%
%INRIA Lille -- Nord, France\and%
%Institut Hospitallo-Universitaire, Strasbourg, France \and%
%University of Lille, Lille, France%
%}
%
\maketitle

\begin{abstract}
The accurate mechanical modeling of liver is of a paramount interest in simulation-based operation planning 
and computer-aided per-operative guidance. However, the realistic simulation of liver behaviour is a challenging task, 
since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, each having different mechanical properties. While the 
mechanical influence of vessels on both global and local behaviour of the organ has been measured and modelled previously~\cite{Peterlik2012}, only local 
effects have been studied in the case of Glisson's capsule~\cite{Hollenstein2006,Ahn2010}.

In this paper we introduce what we believe is the first real-time model of liver taking into account all three constituents, each being simulated 
employing a different finite element formulation.
Using the proposed model, we focus on the role of the Glisson's capsule: first, we show that the model is able to reproduce correctly
the local effects of the capsule in so called aspiration test. Second, based on CT image data of porcine abdominal cavity acquired in supine and flank 
positions, we demonstrate 
that the capsule has also a global impact on the mechanical behaviour of the liver in the case when the organ undergoes large deformations.

%This work provides a complete physical model for the deformations of the liver with its three constituents: parenchyma, vessels and the Glisson's capsule.
%Some recent experimental results \cite{Ahn2010} have shown that there is a significant difference of stiffness between the Glisson's capsule and the parenchyma. 
%Moreover, it has been observed that the capsule plays an important role for the deformation at the local level \cite{Hollenstein2006}.
%However, the Glisson's capsule is very thin and its influence for the global deformations of the liver was never highlighted.
%In this work, we propose to use a membrane FEM model for the capsule, that is coupled with a vascularized model of the liver.
%We show that the approach is able to reproduce the deformations observed at the local level.
%Additionally, we show, by simulating natural global deformations of the liver observed on real images (due to gravity), that the results have important differences with and without the capsule.

\keywords{deformable liver model, Glisson's capsule, vascularization, real-time soft tissue modeling}
\end{abstract}

\section{Introduction} 
According to the statistics, nearly 100,000 European citizens die of cirrhosis or liver cancer each year. 
Although new methods such as radio-frequency- and cryo-ablation known in the interventional radiology 
seem to be promising, surgery remains the option that offers the foremost success rate against these pathologies. 
Nevertheless, surgery is not always performed due to several limitations, in particular the determination 
of accurate eligibility criteria for the patient. 
In this context, the pre-operational planning becomes a crucial task having a significant impact on the treatment.

@@

%Then we show the correct behaviour of the
%model in context of \emph{local} deformations and finally we analyze the influence of
%the capsule on the \emph{global} deformations of the liver and show that despite the
%small thickness it plays an important role also in the \emph{global} context.

%Medical simulation is expected to play an integral part in many aspects of
%medical practice. From diagnosis, through planning or training to computer
%assisted intervention.
%Medical training application, like the ones for training laparoscopic
%surgery, have to operate in real-time and a good medical model has to be
%simple yet efficient to be really usable in this environment.
%
%Liver is largest internal organ formed mostly by a soft tissue (parenchyma)
%that is hyper-elastic and also visco-elastic. The parenchyma is interwoven
%by dense vascular network, this tree-like tubular structure is stiffer then
%the parenchyma. Finally the parenchyma is covered by very stiff collagenous
%membrane that preserves the integrity of the organ.
%
%Strangely, most of the work on liver modelling only considers the
%parenchyma and model the liver as single homogeneous object \TG{Refs}.
%Only recently
%Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
%structures on the liver deformation.
%
%Another overlooked component is the Glisson's capsule.
%Biomechanical measurements report that the capsule has very small
%thickness (approx. 20 $\mu$m for porcine liver \cite{Umale2011}).
%On the other hand it has in three orders of magnitude higher stiffness than
%the parenchyma.
%Because of this it plays an important role in local
%deformations of the liver \cite{Ahn2010,Hollenstein2006}.
%It was reported that evaluation of the liver as a homogeneous material
%without capsule can lead to overestimation of local deformations up to the factor of 3
%\cite{Hollenstein2006}.
%
%While the influence on local deformation was reported, no experimental
%results show whether the capsule also plays an important role in global
%deformations of the liver. We propose a composite model that models the
%capsule with membrane FEM elements. We show that the model is able to
%simulate expected local behaviour reported in literature. Further more we
%show that the capsule plays also an important role in the deformation on
%global scale.

@@

\subsection{Parenchyma and Vascularization} %{{{

It is known that the parenchyma exhibits non-linear viscoelastic behaviour \cite{Marchesseau2010}.
However, as we are mainly interested in the static equilibrium, we do not model the time-dependent
phenomena related to viscosity.

@@

The model of the vascularization is based on linear beams 
with local frames of reference~\cite{Duriez2006}; in many aspects it's similar to the
corotational formulation described above. As such the model also handles geometric
non-linearities in the deformation. Through a specification of cross section and moments of inertia, 
the model can account for the specific properties of the blood vessels.

The beam-based model of vessels is mechanically coupled to the parenchyma as described in~\cite{Peterlik2012}. 
The coupling assumes that there is no relative motion between the vessels and the surrounding parenchyma. 
%The proposed model assumes no relative
%motion of vessels to the parenchyma. Segmented vascular network is
%discretized into the set of linear elements that are then projected into
%the tetrahedral mesh of the parenchyma. 
During the simulation, the nodes of linked beams are first displaced and rotated according to the actual motion of associated tetrahedra. 
As the deformation of beams results in mechanical response represented by forces and torques, these are propagated back to 
the tetrahedral FE model.

%a displacement of tetrahedral nodes are mapped  forces and torques resulting from the deformations of vessels by parenchyma are
%computed and propagated back onto the tetrahedral mesh.

@@

\subsection{Coupling Between Capsule and Parenchyma} %{{{
The literature reports high cohesion between capsule and parenchyma.
Based on this property we assume there is no relative motion of the capsule \wrt\ the parenchyma.
Although an arbitrary surface mesh could be used to model the capsule, we exploit 
the fact that the parenchyma is modeled by tetrahedral elements having
triangular faces. Thus, as the boundary of the volumetric mesh is already
triangulated, we simply employ the triangles on the mesh surface to model the capsule.

@@

\begin{figure}[t]
    	\subfigure[]{\includegraphics[height=3.5cm]{aspiration.jpg}}
	\subfigure[]{\includegraphics[width=7cm]{aspiration.pdf}}
  \caption{\label{fig-aspiration}Aspiration test: (a) the SOFA simulation scene, (b) cube displacement profile
  with the capsule (in blue) and without capsule (in black).}
\end{figure}

@@

@@

\begin{figure}[t]
  \centering
    \includegraphics[width=.95\textwidth]{forceEvolution.pdf}
  \caption{\label{f:forceEvol}Evolution of forces during the simulation, where the supine data is deformed
towards the flank shape, and the stiffness of the capsule is increased from 0 to 8.22\,MPa after the initial equilibrium is achieved.}
\end{figure}

\subsection{Global Deformations}
In the previous section we have shown that the composite model of parenchyma and 
capsule can be used to reproduce the \emph{local} behaviour of the tissue. In the following, 
we first present a real-time complete model of entire liver built from image data. 
Second, using this model we demonstrate the \emph{global} mechanical influence of the capsule when 
the model undergoes large deformations.

@@

\begin{figure}[t]
  \centering
    \subfigure[]{\includegraphics[width=.32\textwidth]{gravity2.png}}
    \subfigure[]{\includegraphics[width=.32\textwidth]{registered.png}}
    \subfigure[]{\includegraphics[width=.32\textwidth]{distanceMap.png}}
    \caption{\label{f:global}%
      (a) Liver under gravity with (red) and without (yellow) capsule;
      (b) liver in supine (yellow) and flank (red) position;
      (c) Hausdorff distance map between registered version with and without capsule.}
\end{figure}

In order to demonstrate the influence of the capsule on a real deformation of the liver, we 
performed a following protocol employing the supine and flank models built from the CT data. 
First, we performed a rigid registration of the two organs \wrt\ the root point of the hepatic vein. 
As shown in Fig.~\ref{f:global}b, the liver undergoes very large deformation when the body 
is turned from the supine to the flank position, exceeding 6\,cm for some parts of the organ. 
In order to induce similar deformation, we extracted feature pairs from the vascular system of the liver:
we selected points such as bifurcations which can be identified easily in both supine and flank 
data resulting in a set of correspondences. In the next step, we used these correspondences as a non-homogeneous
boundary conditions (imposed via penalty term) in order to perform a deformable registration of the supine
model onto the flank data. During the registration, the response forces acting in the points with prescribed displacements
were calculated. It is important to emphasize that the goal of the simulation was not 
to perform an accurate registration of the flank and supine data, but the simulation was used 
as an example of a real deformation of the organ.

The simulation was performed with the model composed of parenchyma,  vessels and capsule, however, zero stiffness was 
set for the last component.
Once the static equilibrium was achieved, the effect of the capsule was gradually switched 
on, \ie\ the stiffness of the membrane was slowly increased from 0 to 8.22\,MPa. The evolution of the force response in the 
points with prescribed displacements is depicted in Fig.~\ref{f:forceEvol} showing the change in the response as the effect
of the capsule is included. At the same time, the shape of the deformed liver with capsule was compared 
to the model without the capsule. A Hausdorff distance map for those cases is depicted in Fig.~\ref{f:global}c: in the red areas, the 
maximal difference exceeding 1.25\,cm was measured. We believe that this result confirms that the Glisson's capsule 
plays an important role in the global mechanical behaviour of the liver.

@@

%
% ---- Bibliography ----
%

\bibliographystyle{splncs03}
\bibliography{bibdata}

>>>>

%\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
\documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
\usepackage{subfig}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:
%%
%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}
% \biboptions{}

\def\wrt{w.\,r.\,t.\ }
\def\eg{e.\,g.\ }
\def\ie{i.\,e.\ }
\def\Dash{\nobreak\,---\penalty-500\,}

@@

\renewcommand{\Vec}[1]{\boldsymbol{#1}}
\newcommand{\Mat}[1]{\boldsymbol{#1}}

\journal{Medical Image Analysis}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{TODO: Title} % TODO

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author{}

\address{}

\begin{abstract}
%% Text of abstract
TODO % TODO
\end{abstract}

\begin{keyword}
deformable liver model \sep Glisson's capsule \sep vascularization \sep real-time soft tissue modeling

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

% ---  Main text  ---

\TG{mostly just copy of MICCAI paper}

\section{Introduction} %{{{

According to the statistics, nearly 100,000 European citizens die of cirrhosis or liver cancer each year. 
Although new methods such as radio-frequency- and cryo-ablation known in the interventional radiology 
seem to be promising, surgery remains the option that offers the foremost success rate against these pathologies. 
Nevertheless, surgery is not always performed due to several limitations, in particular the determination 
of accurate eligibility criteria for the patient. 
In this context, the pre-operational planning becomes a crucial task having a significant impact on the treatment.

@@

@@

\subsection{Parenchyma} %{{{

It is known that the parenchyma exhibits non-linear viscoelastic behaviour
\cite{Marchesseau2010,Gao2009}.
However, as we are mainly interested in the static equilibrium, we do not model the time-dependent
phenomena related to viscosity.

@@

\subsection{Vascularization} %{{{

As a first step towards a mechanical model of vessels, a continuous
representation of vascular structures is constructed. Using segmeted data
of a vessel tere, points along centerlines of each vessel branch are
selected. This task can be performed either manually or automatically, \eg
with VMTK\footnote{www.vmtk.org}. Each branch can be represented as a
series of cubic B\'{e}zier curves fitted to the centerline points: the
series begins in a starting point of the branch and ends either in another
branching point or in the endpoint of the branch.

A set of interpolated points is constructed using the B\'{e}zier curves;
positions of the interpolated points can be chosen arbitrarily along each
branch dividing it ionto intervals with a constant or varying length. For
each interpolated point we calculate its orientation given by the tangent
and normal of the B\'{e}zier curve in that point.

The interpolated points are used as nodes of serially linked beam elements,
in a similar way as proposed by Duriez et al. \cite{Duriez2006} for
simulating catheters and guidewires. This model shares some similarities
with the corotational model described above, and in particular allows for
geometrically non-linear deformations. We introduce some modifications to
the model to take into account particular nature of vessels, in particualr
through specific cross section profiles and moments of inertia (see
\cite{Przemieniecki1985} for details). The static formulation for the
deformation of a beam is described by a system similar to the one used for
the parenchyme, with the difference that each node is described with six
degrees of freedom, three of which correspond to the spatial position, and
three to the angular position of the node in a global reference frame
\TODO{figure?}.

% TODO: presunout do sekce composite model
%The beam-based model of vessels is mechanically coupled to the parenchyma as described in~\cite{Peterlik2012}. 
%The coupling assumes that there is no relative motion between the vessels and the surrounding parenchyma. 
%%The proposed model assumes no relative
%%motion of vessels to the parenchyma. Segmented vascular network is
%%discretized into the set of linear elements that are then projected into
%%the tetrahedral mesh of the parenchyma. 
%During the simulation, the nodes of linked beams are first displaced and rotated according to the actual motion of associated tetrahedra. 
%As the deformation of beams results in mechanical response represented by forces and torques, these are propagated back to 
%the tetrahedral FE model.

%%a displacement of tetrahedral nodes are mapped  forces and torques resulting from the deformations of vessels by parenchyma are
%%computed and propagated back onto the tetrahedral mesh.

@@

\subsection{Coupling Between Vessels and Parenchyma} %{{{

We propose a mapping betwen the mesh nodes of the vessels and parenchyma to
create the vascularized model. Since no relative motion between the vessels
and parenchyma is observed in reality, the mapping betweeen the two can be
modelled as a constraint. In each step of the simulation, the actual
displacements of the parenchyma mesh nodes are mapped to the vessel nodes
and reciprocally, the force contribution due to the deformation of the
vessel is propagated to the parenchyma. The mapping of the forces is based
on a principle of virtual work. Using the corotational model, the
parenchyma is discretized by a mesh composed of a linear $P_1$ tetrahedral
elements given by four nodes with three degrees of freedom. The vessel is
modeled with beam elements where each beam is given by two end points each
having 6 degreees of freedom (positions and rotations). We focus on mapping
betwen an arbitrary beam node $b$ and tetrahedron $T$.

Let us denote $\Vec{p}$ the initial position and $\Mat{R}$ initial rotation
of the beam node $b$. Before the simulation starts, we select the
tetrahedron $T = \{\Vec{t}_A, \Vec{t}_B, \Vec{t}_C, \Vec{t}_D \}$ which is
closest to the beam node $b$. Then, the barycentric coordinates
$\Vec{\beta}_T = \{\beta_A, \beta_B, \beta_C\}$ of the position $\Vec{p}$
are computed \wrt the tetrahedron $T$ and stored. Similarly, the initial
rotation $\Mat{R}$ is transformed to a a matrix $\Mat{O}_T$ which defines
the orientation of the point $B$ \wrt the local basis $\Mat{L}$ defined by
the edges of the tetrahedron, \ie $\Mat{O}_T = \Mat{L}^{-1} \Mat{R}$.
According to the initial assumption (no relative motion between parenchyma
and vessel), both the barycentric coordinates $\Vec{\beta}_T$ and
$\Mat{O}_T$ are constant during the simulation. Let's suppose that in  the
actual step, the tetrahedron is rotated and deformed as shown in
\TODO{figure}. Denoting the nodal positions of the deformed tetrahedron $T'
= \{\Vec{t}'_A, \Vec{t}'_B, \Vec{t}'_C, \Vec{t}'_D\}$, new actual position
$\Vec{p}'$  of the beam node is computed using the barycentric coordinates
$\Vec{\beta}_T$ as
%
\begin{equation}
  \Vec{p}' = \Vec{t}'_A \beta_A + \Vec{t}'_B \beta_B + \Vec{t}'_C \beta_C +
  \Vec{t}'_D (1 - \beta_A - \beta_B - \beta_C).
\end{equation}
%
The updated orientation $\Mat{R}'$ of the beam node is computed using the
basis $\Mat{L}'$ given by the actual displacement of the tetrahedron nodes
and precomputed orientation $\Mat{O}_T$. First, $\Mat{\tilde{R}}' =
\Mat{L}' \Mat{O}_T$ is computed. However, since besides the rotation ad
deformation of the tetrahedron must be taken into account the matrix
$\Mat{\tilde{R}}'$ is not necessarily orthogonal. Therefore, the updated
orientation $\Mat{R}'$ is calculated as orthogonal component of polar
decomposition of $\Mat{\tilde{R}}'$. As the positional mapping is applied
in each step of the simulation, the positions of the beams are
kinematically linked to the positions of tetrahedra.

After the updated position is mapped from the tetrahedron to the beam point,
the forces acting in the beam point must be mapped onto the vertices of the
associated tetrahedron. For the beam point, we have linear forces $\Vec{f}
= \{ f_x , f_y , f_z \}$
as well as torques $\Vec{\tau} = \{ \tau_x, \tau_y, \tau_z \}$, whereas for the tetrahedron, there are only
linear nodal forces $\Mat{F} = \{ \Vec{f}_A , \Vec{f}_B , \Vec{f}_C ,
\Vec{f}_D \}$ . The force contribution to each node
given by linear forces f is computed using the barycentric coordinates, required
by the positional mapping. The torques are transformed to linear forces acting in
the tetrahedron nodes using the equation $\Vec{\tau} = \Vec{r} \times
\Vec{f}$ where $\Vec{r}$ is vector connecting
the beam and tetrahedron nodes. Putting it together, the force in tetrahedron
node $I \in \{A, B, C, D\}$ is computed as $\Vec{F}_I = \beta_I \Vec{f} −
(\Vec{t}'_I − \Vec{p}') \times \tau$.

\subsection{Coupling Between Capsule and Parenchyma} %{{{
The literature reports high cohesion between capsule and parenchyma
\TODO{need refs}.
Based on this property we assume there is no relative motion of the capsule \wrt the parenchyma.
Although an arbitrary surface mesh could be used to model the capsule, we exploit 
the fact that the parenchyma is modeled by tetrahedral elements having
triangular faces. Thus, as the boundary of the volumetric mesh is already
triangulated, we simply employ the triangles on the mesh surface to model the capsule.

@@

%XXX
%\begin{figure}[t]
%    	\subfigure[]{\includegraphics[height=3.5cm]{aspiration.jpg}}
%	\subfigure[]{\includegraphics[width=7cm]{aspiration.pdf}}
%  \caption{\label{fig-aspiration}Aspiration test: (a) the SOFA simulation scene, (b) cube displacement profile
%  with the capsule (in blue) and without capsule (in black).}
%\end{figure}

@@

\TG{
  We are able to match the deformations presented in
  [Hollenstein2006] with less then 1\% error using the following
  parameters: E=30\,kPa for parenchyma; E=3\,MPa and t=93\,$\mu$mm for capsule.
  The paper does not specify single values for elasticity modulus but specifies
  the possible range. Our values lie in the prescribed ranges.}

@@

%XXX
%\begin{figure}[t]
%  \centering
%    \includegraphics[width=.95\textwidth]{forceEvolution.pdf}
%  \caption{\label{f:forceEvol}Evolution of forces during the simulation, where the supine data is deformed
%towards the flank shape, and the stiffness of the capsule is increased from 0 to 8.22\,MPa after the initial equilibrium is achieved.}
%\end{figure}

\subsection{Global Deformations} %{{{

In the previous section we have shown that the composite model of parenchyma and 
capsule can be used to reproduce the \emph{local} behaviour of the tissue. In the following, 
we first present a real-time complete model of entire liver built from image data. 
Second, using this model we demonstrate the \emph{global} mechanical influence of the capsule when 
the model undergoes large deformations.

@@

%XXX
%\begin{figure}[t]
%  \centering
%    \subfigure[]{\includegraphics[width=.32\textwidth]{gravity2.png}}
%    \subfigure[]{\includegraphics[width=.32\textwidth]{registered.png}}
%    \subfigure[]{\includegraphics[width=.32\textwidth]{distanceMap.png}}
%    \caption{\label{f:global}%
%      (a) Liver under gravity with (red) and without (yellow) capsule;
%      (b) liver in supine (yellow) and flank (red) position;
%      (c) Hausdorff distance map between registered version with and without capsule.}
%\end{figure}

In order to demonstrate the influence of the capsule on a real deformation of the liver, we 
performed a following protocol employing the supine and flank models built from the CT data. 
First, we performed a rigid registration of the two organs \wrt the root point of the hepatic vein. 
As shown in Fig.~\ref{f:global}b, the liver undergoes very large deformation when the body 
is turned from the supine to the flank position, exceeding 6\,cm for some parts of the organ. 
In order to induce similar deformation, we extracted feature pairs from the vascular system of the liver:
we selected points such as bifurcations which can be identified easily in both supine and flank 
data resulting in a set of correspondences. In the next step, we used these correspondences as a non-homogeneous
boundary conditions (imposed via penalty term) in order to perform a deformable registration of the supine
model onto the flank data. During the registration, the response forces acting in the points with prescribed displacements
were calculated. It is important to emphasize that the goal of the simulation was not 
to perform an accurate registration of the flank and supine data, but the simulation was used 
as an example of a real deformation of the organ.

The simulation was performed with the model composed of parenchyma,  vessels and capsule, however, zero stiffness was 
set for the last component.
Once the static equilibrium was achieved, the effect of the capsule was gradually switched 
on, \ie the stiffness of the membrane was slowly increased from 0 to 8.22\,MPa. The evolution of the force response in the 
points with prescribed displacements is depicted in Fig.~\ref{f:forceEvol} showing the change in the response as the effect
of the capsule is included. At the same time, the shape of the deformed liver with capsule was compared 
to the model without the capsule. A Hausdorff distance map for those cases is depicted in Fig.~\ref{f:global}c: in the red areas, the 
maximal difference exceeding 1.25\,cm was measured. We believe that this result confirms that the Glisson's capsule 
plays an important role in the global mechanical behaviour of the liver.

\subsection{More Experiments}

\TG{list of other experiments}

@@

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

% ---  Bibliography  ---

\bibliographystyle{model1-num-names}
\bibliography{bibdata}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use model1-num-names.bst.
